### 2016-06-13 Projects/DelCo_2014_2015/Reports/Hose_Stream_Tests/TeX_Files/DelCo_Hose_Stream_Tests.tex
Resolved merge issues with last two versions of hose stream report

% ================
% = Introduction =
% ================
\chapter{Introduction}
\label{chap:intro}
\section{Background}
\label{sec:background}
% More hose stream studies and explanations

One of the objectives of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to improve the safety and effectiveness of firefighters through improved knowledge of fire behavior and examine how different firefighting tactics affect the fire environment. NIST and UL FSRI have conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire~\cite{madrzykowski2009fire,kerber2009fire}. The studies have resulted in guidance to the fire service regarding ventilation tactics. However, ventilation tactics alone will not result in the extinguishment of the fire; fire suppression with hose streams is also needed.

Fire suppression tactics using hose streams also affect the ventilation in a structure and may impact the movement of smoke and heat through a structure. As vents are made to advance the hoseline or if ventilation inducing handline tactics are in practice, hose stream impacts on ventilation may be even more significant. Additional research addressing the coordination of suppression tactics and the impact on ventilation is needed to complete recommendations on fire control tactics to appropriate standards, education, and training documents.

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is an adjustable-pattern spray nozzle, commonly referred to as a combination nozzle~\cite{NFPA_1964}. Most combination nozzles contain an adjustable tip that can be rotated to produce different stream patterns. The types of streams produced by a combination nozzle are often divided into three categories based on the angle of the stream with respect to the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}; a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}; and a wide fog stream produces an angle between 45\SI{}{\degree} and 80\SI{}{\degree}~\cite{IFSTA:Essentials_of_FF}. Increasing the angle of the stream allows for a larger area to be covered by the water spray.  However, increasing the angle decreases the distance reached by the stream and its penetration power. 
%\textit{Essentials of Fire Fighting} defines 
%https://docs.google.com/presentation/d/1hZ5DZhYXxCkcIEHNZKumH48b_Q53iM2D7asXkMYlVMM/embed?hl=en&size=l&slide=id.p39

%http://www.tft.com/literature/library/files/bfa.pdf
%Keith Royer, "Water for Fire Fighting", (Engineering Extension Service, Iowa State University, Ames, Iowa) p. 6 
In addition to the type of stream pattern used, the method or motion of the nozzle used to apply the hose stream can also vary. For example, Royer suggested using a clockwise rotation instead of a counterclockwise nozzle movement pattern for the following reasons~\cite{Royer:ISU}:
\begin{enumerate} 
	\item Clockwise rotation drives most of the heated gases, smoke, and flames away from the nozzle, while counterclockwise rotation does just the opposite.
	\item Steam formed by the clockwise rotation has a violent rolling action. Counterclockwise rotation produces steam with an inactive and lazy action.
	\item A clockwise rotation increases the efficiency of water and produces a faster knockdown time than counterclockwise rotation.
\end{enumerate}

\section{Objectives}
\label{sec:objectives}
This report is concerned with examining the impact of hose stream and application pattern, or nozzle motion, selection on air movement inside two full scale test structures built to represent residential sized structures. Numerous experiments containing different flow path configurations were conducted to study the impact of hose stream and nozzle movement selection on ventilation in a structure. More specifically, the experiments were conducted to fulfill the following objectives:
\begin{enumerate}
	\item Determine which hose stream patterns and nozzle movement patterns cause significant air movement in a structure 
	\item Determine the amount of air movement induced by different hose streams and nozzle movement patterns

\end{enumerate}
Four types of hose streams and four different nozzle movement patterns were studied throughout the experiments described in this report. Three types of hose stream patterns using a combination nozzle were studied: straight stream, narrow fog stream, and wide fog stream. Water was flowed into the structure by having the hose in a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Additionally, a set of experiments comparing the impact of a straight stream from a combination nozzle and a solid stream from a smooth bore nozzle equipped with a 1~in tip on air flows in a structure was conducted. Images of the four types of hose streams are presented in Fig.~\ref{fig:hose_streams}.

>>>>

% ================
% = Introduction =
% ================
\chapter{Introduction}
\label{chap:intro}
\section{Background}
\label{sec:background}
One of the objectives of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to improve the safety and effectiveness of firefighters through improved knowledge of fire behavior and examination of how different firefighting tactics affect the fire environment. NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire~\cite{madrzykowski2009fire,kerber2009fire}. The studies have resulted in guidance to the fire service regarding ventilation tactics. However, ventilation tactics alone will not extinguish the fire; fire suppression with hose streams is also needed.

Fire suppression tactics using hose streams also affect the ventilation in a structure and may impact the movement of smoke and heat through a structure. If vents are made to advance the hoseline or if ventilation inducing handline tactics are in practice, hose stream impacts on ventilation may be even more significant. Additional research addressing the coordination of suppression tactics and the impact on ventilation is needed to complete recommendations on fire control tactics to appropriate standards, education, and training documents.

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is an adjustable-pattern spray nozzle, commonly referred to as a combination nozzle~\cite{NFPA_1964}. Most combination nozzles contain an adjustable tip that can be rotated to produce different stream patterns. The types of streams produced by a combination nozzle are often divided into three categories based on the angle of the stream with respect to the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}; a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}; and a wide fog stream produces an angle between 45\SI{}{\degree} and 80\SI{}{\degree}~\cite{IFSTA:Essentials_of_FF}. Increasing the angle of the stream allows for a larger area to be covered by the water spray.  However, increasing the angle decreases the distance reached by the stream and its penetration power.

In addition to the type of stream pattern used, the method or motion of the nozzle used to apply the hose stream can also vary. For example, Royer suggested using a clockwise rotation instead of a counterclockwise nozzle movement pattern for the following reasons~\cite{Royer:ISU}:
\begin{enumerate} 
	\item Clockwise rotation drives most of the heated gases, smoke, and flames away from the nozzle, while counterclockwise rotation does just the opposite.
	\item Steam formed by the clockwise rotation has a violent rolling action. Counterclockwise rotation produces steam with an inactive and lazy action.
	\item A clockwise rotation increases the efficiency of water and produces a faster knockdown time than counterclockwise rotation.
\end{enumerate}

\section{Objectives}
\label{sec:objectives}
This report focuses on examining the impact of hose stream and application pattern (or nozzle motion) selection on air movement inside two full-scale test structures built to represent residential sized structures. Experiments containing different flow path configurations were conducted to study the impact of hose stream and water application selection on ventilation in a structure. More specifically, the experiments were conducted to fulfill the following objectives:

\begin{enumerate}
	\item Determine which hose stream patterns and nozzle movement patterns cause significant air movement in a structure 
	\item Determine the amount of air movement induced by different hose streams and nozzle movement patterns

\end{enumerate}

Four types of hose streams and four nozzle movement patterns patterns were studied as part of the experiments described in this report. Three of the hose stream patterns studied were from a combination nozzle: straight stream, narrow fog stream, and wide fog stream. Water was flowed into the structure by having the hose in a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Additionally, experiments comparing the differences in impact on air flows in a structure from a combination nozzle straight stream and a solid stream from a smooth bore nozzle equipped with a 1~in tip were conducted. Images of the four types of hose streams are presented in Fig.~\ref{fig:hose_streams}.
